\sysfull

### 2017-11-16 paper/intro.tex
intro revised

\textbf{Task Decomposition of Intermittent Programs.} Recently proposed {\em
task-based} programming and execution models~\cite{alpaca,chain} advocate for
static decomposition of a program (by a programmer) into a collection of tasks.
Tasks can include arbitrary computation and, upon completion, are guaranteed to
have executed {\em atomically}, despite arbitrarily-timed power failures.
The programmer explicitly expresses task-to-task control flow. The execution
in Figure~\ref{fig:coalesce} (top) illustrates how the decomposition into
tasks affects the time the application takes to complete. Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory by the task, to maintain consistency of program state~\cite{chain,alpaca}. Naively minimizing the number of task transitions, (i.e. the task count), \emph{at compile time} risks creating a task that requires more energy than the device can buffer in its storage capacitor. Such a large task would fail to complete unless the environment provides
sufficient incoming energy to supplement the energy in the capacitor. To eliminate this risk, the programmer (or the compiler) must decompose the program conservatively into many small tasks, at the cost of \emph{higher task transition overhead}.

One approach to decrease the task count, without assuming a minimum level of
incoming power always available, is to predict task energy consumption and
choose a decomposition specific to the device capacitor size. However, statically predicting energy of arbitrary input-dependent code with peripheral access is a problem without a general solution. Furthermore, a static decomposition approach cannot produce decompositions portable across devices with different storage capacitors.

@@

\textbf{Task Coalescing.} To address this challenge we propose a
\emph{dynamic} task decomposition approach, that accepts any static
decomposition and opportunistically coalesces tasks at runtime when
there is sufficient energy. \emph{Task coalescing} is a mechanism where
atomic tasks are merged into a single (larger) task while preserving
atomicity of the merged task. If power fails while a colesced task is
running, execution will restart at the first task in the group.
Figure~\ref{fig:coalescing} (bottom) illustrates how dynamic coalescing
changes the earlier execution and highlights the change in memory model
necessary to not break atomicity.

@@

Task coalescing removes the optimization burden of choosing the best task decomposition from the programmer, because it accepts any decomposition and improves it dynamically. To further reduce the programmer effort, we propose a compiler pass for \emph{automatic decomposition} of programs into (small) \emph{atomic} tasks. The compiler identifies non-volatile variables shared across tasks as tasks are created, and instruments reads and writes of those variables, using memory virtualization, to keep the data consistent in the presence of power loss. Despite being limited to a subset of the C language, the automatic task decomposition allowed us to port several
applications to an intermittent platform with a moderate effort.

\textbf{Contributions.} We develop \sys\footnote{To be released at \url{http://anonymized.link}}: a task coalescing system for task-based programming models on intermittently-powered devices. The capabilities required for effective coalescing mark our contributions:

>>>>

% <<<<<<< HEAD
% \begin{figure}
% 	\includegraphics[width=.5\columnwidth]{figures/alpaca-war.pdf}
% 	\caption{\label{fig:virtualization}Adaptive task division at run-time might lead to write-after-read dependencies (variable \ttfamily{x}) that do not exist with static task division.\todo{Box with Task 1,2,3 must be the same size as merged task; Merged task must be named as Task 4; x must be marked in text as having WAR; "Task X" name should be next to the task box; "Window" - small letter}{Sinan}}
% \end{figure}
% =======

\textbf{Task Decomposition of Intermittent Programs.} Recently proposed {\em
task-based} programming and execution models~\cite{alpaca,chain} advocate for
static decomposition of a program (by a programmer) into a collection of tasks.
Tasks can include arbitrary computation and, upon completion, are guaranteed to
have executed {\em atomically}, despite arbitrarily-timed power failures.
The programmer explicitly expresses task-to-task control flow. The execution
in Figure~\ref{fig:coalesce} (top) illustrates how the decomposition into
tasks affects the time the application takes to complete. Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}. Naively minimizing the number of task transitions, (i.e. the task count), \emph{at compile time} risks creating a task that requires more energy than the device can buffer in its energy storage. Such a large task would fail to complete unless the environment provides sufficient incoming energy to supplement the energy in the capacitor. To eliminate this risk, the programmer (or the compiler) must decompose the program conservatively into many small tasks, at the cost of \emph{higher task transition overhead}.

One approach to decrease the task count, without assuming a minimum level of
incoming power always available, is to predict task energy consumption and
choose a decomposition specific to the device capacitor size. However, statically predicting energy of arbitrary input-dependent code with peripheral access is a problem without a general solution. Furthermore, a static decomposition approach prevents portability across devices with different storage capacitors.

@@

\textbf{Task Coalescing.} To address this challenge we propose a \emph{dynamic} task decomposition, that accepts any static decomposition and opportunistically coalesces tasks at runtime when there is sufficient energy. With \emph{task coalescing} atomic tasks are merged into a single (larger) task while preserving atomicity of the merged task. If power fails while a coalesced task is running, execution will restart at the first task in the group. Figure~\ref{fig:coalescing} (bottom) illustrates how dynamic coalescing changes the earlier execution and highlights the change in memory model necessary to not break atomicity.

@@

Task coalescing removes the burden from the programmer, because it accepts any task decomposition and improves it dynamically. To further reduce the programmer effort, we propose a compiler pass for \emph{automatic decomposition} of programs into (small) \emph{atomic} tasks. The compiler identifies non-volatile variables shared across tasks as tasks are created, and instruments reads and writes of those variables, using memory virtualization, to keep the data consistent in the presence of power loss. Despite being limited to a subset of the C language, the automatic task decomposition allowed us to port several
applications to an intermittent platform with a moderate effort.

% <<<<<<< HEAD
% We develop \sys%
% \footnote{To be released at \url{http://anonymized.link}}
% %
% : a task coallescing system for task-based
% programming models on intermittently-powered devices. The
% capabilities required for effective coallescing mark our
% contributions:
% >>>>>>> 96e5d9f9b4e97ce8700c375bd8e11708870f5aba
% =======
\textbf{Contributions.} We develop \sys\footnote{To be released at \url{http://anonymized.link}}: a task coalescing system for task-based programming models on intermittently-powered devices. The capabilities required for effective coalescing mark our contributions:
% >>>>>>> 84d7711b6c6f6424cdad97efc8aad96e52f316ff
